Actuator line modeling of vertical-axis turbines

### 2018-03-29 reviews/reviewer2-response.md

# Response to reviewer 2

We thank the reviewer for his or her insightful comments, and feel that these
have helped improve the manuscript significantly.

### 2018-03-29 paper.tex
Add revised preprint date
Fix typo
Switch to Energies format

% wedoc.tex V1.0, 17 June 2010

\documentclass[times]{weauth}

\usepackage{moreverb}

\usepackage{lineno}

\usepackage{nomencl}
\makenomenclature

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

@@

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2016}

\begin{document}

\runningheads{P.~Bachant \emph{et al.}}{Actuator line modeling of vertical-axis
turbines}

\articletype{RESEARCH ARTICLE PREPRINT (submitted May 2016)}

\title{Actuator line modeling of vertical-axis turbines}

\author{P.~Bachant\textsuperscript{1}, A.~Goude\textsuperscript{2}, and
M.~Wosnik\textsuperscript{1}}

\address{\textsuperscript{1}Center for Ocean Renewable Energy,
University of New Hampshire, 24
Colovos Rd., Durham, NH, 03824, USA \\
\textsuperscript{2}Department of Engineering Sciences,
Division of Electricity, Uppsala University, Uppsala 751 21, Sweden}

\corraddr{P.~Bachant, Chase Ocean Engineering Laboratory, University of New
Hampshire, 24 Colovos Rd., Durham, NH, 03824, USA. \\
E-mail: pwbachant@gmail.com}

\linenumbers

\begin{abstract}
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a combination of classical blade element theory and
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested. The RANS models were able to
    be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
\end{abstract}

\keywords{wind turbine; VAWT; cross-flow turbine; blade element theory; dynamic
stall; lifting line}

\maketitle

\printnomenclature

@@

\nomenclature{CFT}{Cross-flow turbine}
\nomenclature{VAT}{Vertical-axis turbine}
\nomenclature{AFT}{Axial-flow turbine}
\nomenclature{HAT}{Horizontal-axis turbine}

@@

\nomenclature{RANS}{Reynolds-averaged Navier--Stokes}

@@

\nomenclature{ALM}{Actuator line model}
\nomenclature{LES}{Large eddy simulation}

@@

\nomenclature{$c$}{Blade chord length}
\nomenclature{$R$}{Maximum rotor radius}
\nomenclature{$Re_D$}{Rotor diameter Reynolds number}
\nomenclature{DOE}{US Department of Energy}
\nomenclature{SNL}{Sandia National Laboratories}

@@

\nomenclature{$\vec{U}_\mathrm{rel}$}{Blade element local relative velocity}
\nomenclature{$\alpha$}{Blade element angle of attack}
\nomenclature{$\omega$}{Rotor angular velocity}
\nomenclature{$r$}{Blade element radius}
\nomenclature{$\vec{U}_\mathrm{in}$}{Blade element local inflow velocity}

@@

\nomenclature{$U_\infty$}{Free stream velocity}

@@

\nomenclature{$\lambda$}{Rotor tip speed ratio}
\nomenclature{$k$}{Reduced frequency or turbulence kinetic energy}

@@

\nomenclature{$C_l$}{Blade element lift coefficient}
\nomenclature{$C_d$}{Blade element drag coefficient}
\nomenclature{$C_m$}{Blade element moment coefficient}

@@

\nomenclature{$Re_c$}{Blade chord Reynolds number}

@@

    \caption{Computational domain used for the RANS ALM simulations. All
    dimensions are in meters, flow is oriented in the $+x$-direction, and the
    cylindrical region represents the swept area of the UNH-RVAT rotor. Note
    that increased mesh refinrement was used for the LES simulations.}

@@

\nomenclature{$\epsilon$}{Gaussion force smoothing function width}

@@

\nomenclature{$C_P$}{Rotor power coefficient}
\nomenclature{$C_D$}{Overall rotor drag (a.k.a. thrust) coefficient}
\nomenclature{$D$}{Rotor diameter}

@@

\nomenclature{$H$}{Rotor height}

@@

\nomenclature{$\lambda_0$}{Tip speed ratio of maximum power coefficient}

@@

\acks The authors would like to acknowledge funding through a National Science
Foundation CAREER award (principal investigator Martin Wosnik, NSF 1150797,
Energy for Sustainability, program manager Gregory L. Rorrer). This work has
been carried out as part of author P. Bachant's doctoral research at the
University of New Hampshire.

\bibliographystyle{wileyj}
\bibliography{library}
\end{document}

>>>>

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g. \documentclass[preprints,article,accept,moreauthors,pdftex,10pt,a4paper]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, cleantechnol, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochem, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicina, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, neuroglia, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, polysaccharides, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, reactions, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{1}
\pubyear{2018}
\copyrightyear{2018}
%\externaleditor{Academic Editor: name}
\history{
    Preprint uploaded XXXXXX
    Received: date;
    Accepted: date;
    Published: date
}

%\updates{yes} % If there is an update available, un-comment this line

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz, siunitx

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Actuator line modeling of vertical-axis turbines}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0003-4039-6954} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{
    Peter Bachant $^{1,\dagger,\ddagger}$\orcidA{},
    Anders Goude $^{2,\ddagger}$
    and Martin Wosnik $^{1,}$*
}

% Authors, for metadata in PDF
\AuthorNames{Peter Bachant, Anders Goude and Martin Wosnik}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad University of New Hampshire, USA\\
$^{2}$ \quad Uppsala University, Sweden}

% Contact information of the corresponding author
\corres{Correspondence: pwbachant@gmail.com}

% Current address and/or shared authorship
% \firstnote{Current address: Affiliation 3}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
%\simplesumm{}

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a combination of classical blade element theory and
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested. The RANS models were able to
    be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
}

% Keywords
\keyword{wind turbine; VAWT; cross-flow turbine; blade element theory; dynamic
stall; lifting line}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

@@

% \nomenclature{CFT}{Cross-flow turbine}
% \nomenclature{VAT}{Vertical-axis turbine}
% \nomenclature{AFT}{Axial-flow turbine}
% \nomenclature{HAT}{Horizontal-axis turbine}

@@

% \nomenclature{RANS}{Reynolds-averaged Navier--Stokes}

@@

% \nomenclature{ALM}{Actuator line model}
% \nomenclature{LES}{Large eddy simulation}

@@

% \nomenclature{$c$}{Blade chord length}
% \nomenclature{$R$}{Maximum rotor radius}
% \nomenclature{$Re_D$}{Rotor diameter Reynolds number}
% \nomenclature{DOE}{US Department of Energy}
% \nomenclature{SNL}{Sandia National Laboratories}

@@

% \nomenclature{$\vec{U}_\mathrm{rel}$}{Blade element local relative velocity}
% \nomenclature{$\alpha$}{Blade element angle of attack}
% \nomenclature{$\omega$}{Rotor angular velocity}
% \nomenclature{$r$}{Blade element radius}
% \nomenclature{$\vec{U}_\mathrm{in}$}{Blade element local inflow velocity}

@@

% \nomenclature{$U_\infty$}{Free stream velocity}

@@

% \nomenclature{$\lambda$}{Rotor tip speed ratio}
% \nomenclature{$k$}{Reduced frequency or turbulence kinetic energy}

@@

% \nomenclature{$C_l$}{Blade element lift coefficient}
% \nomenclature{$C_d$}{Blade element drag coefficient}
% \nomenclature{$C_m$}{Blade element moment coefficient}

@@

% \nomenclature{$Re_c$}{Blade chord Reynolds number}

@@

    \caption{Computational domain used for the RANS ALM simulations. All
    dimensions are in meters, flow is oriented in the $+x$-direction, and the
    cylindrical region represents the swept area of the UNH-RVAT rotor. Note
    that increased mesh refinement was used for the LES simulations.}

@@

% \nomenclature{$\epsilon$}{Gaussion force smoothing function width}

@@

% \nomenclature{$C_P$}{Rotor power coefficient}
% \nomenclature{$C_D$}{Overall rotor drag (a.k.a. thrust) coefficient}
% \nomenclature{$D$}{Rotor diameter}

@@

% \nomenclature{$H$}{Rotor height}

@@

% \nomenclature{$\lambda_0$}{Tip speed ratio of maximum power coefficient}

@@

\acknowledgments{
    The authors would like to acknowledge funding through a National Science
    Foundation CAREER award (principal investigator Martin Wosnik, NSF 1150797,
    Energy for Sustainability, program manager Gregory L. Rorrer). This work has
    been carried out as part of author P. Bachant's doctoral research at the
    University of New Hampshire.
}

\authorcontributions{
    P.B. wrote the computer code, ran the simulations, post-processed the
    results, and wrote the paper.
    A.G. contributed to the computer code and the manuscript.
    M.W. contributed to writing and editing of the manuscript.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflicts of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{library}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
